---
title: "Displaying verbatim code chunks in RMarkdown and Xaringan presentations"
description: |
  Because the best way to teach RMarkdown is with RMarkdown.
author:
  - name: Thomas Mock
    url: https://twitter.com/thomas_mock
date: 08-27-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 4
preview: preview.png
twitter:
  site: "@thomas_mock"
  creator: "@thomas_mock"
categories:
  - meta
  - xaringan
  - rmarkdown
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
xaringanExtra::use_panelset()
```

This will be a short article, but I wanted to note it for my own future use and hopefully a discoverable resource if folks get stuck trying to do the same thing!

I recently gave a presentation on Advanced RMarkdown use, and part of what I was displaying in my RMarkdown generated `{xaringan}` slides was literal RMarkdown code chunks. These verbatim code chunks are very useful and in some cases *necessary* to accurately convey a teaching example.

<iframe width="560" height="400" src="https://www.youtube.com/embed/WkF7nqEYF1E" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

---

Notably, while these verbatim code chunks are extremely useful some difficulty can arise as you're essentially nesting R and RMarkdown code/logic _inside_ a RMarkdown document. In order to "tell" RMarkdown to ignore the code and simply print it out verbatim you need to use a few different techniques.

More details can be found in the [RMarkdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/verbatim-code-chunks.html).

## Verbatim Code Chunks

The primary problem here is we need to "tell" RMarkdown not to parse the RMarkdown code chunk as _real_ code. So we need to account for the 3x backticks that make up an R code chunk as seen below.

```{r, echo = FALSE}
cat("```{r, echo = TRUE}
...code_goes_here...
```")
```

### `cat()` to the rescue

You can use `cat()` to output specific code chunks verbatim as raw text. This can be very fast, but requires you to write code by hand in text.

```{r echo = TRUE, comment = NA}

output_code <-
"````
```{r, eval=TRUE}`r ''`
library(dplyr)
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg))
```\n````"

cat(output_code)
```


### We need more backticks captain!

To return unevaluated verbatim code chunks code you need N + 1 of the max backticks, so for RMarkdown chunks that have 3x backticks, you need to wrap it in 4x backticks on either side of the code.

You _also_ need to invalidate the code chunk by adding a blank `r knitr::inline_expr("''")` to the _end_ of the existing code chunk (ie immediately after the `{r}` like so:

```{r, echo = FALSE}
cat("```{r}`r''`
...code_goes_here...
```")
```

As a code example I can wrap my code in 4x backticks around the 3x backticks, also note the invalidated `r` expression after the `{r}`.

```{r, echo = FALSE}

code_out <- "````
```{r}`r ''`
library(dplyr)
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg))
```\n````"
cat(code_out)
```

That rendered inside RMarkdown outputs:

````
```{r}`r ''`
library(dplyr)
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg))
```

````

However you may notice that there's no grey `code` background for the code chunk. To format it, we can also add the code type to the end of the 4x backticks for formatting purposes. I have a panelset of some of the various code types appended to the 4x backticks below. This affects some highlighting of specific functions/strings/etc in the code as well.

A note here is that my blog is written in `{distill}`, so what you see here is not identical to `{xaringan}` which is built on `remark.js`. As such, I've included the `{distill}` output below, and after that another embedded version of a minimal `{xaringan}` presentation. In `{xaringan}` there is also the `highlightStyle` argument in the YAML which allows for a lot of customization of the code chunk/code/comments colors. The various options for this argument are at the bottom of the [`remark.js` wiki](https://github.com/gnab/remark/wiki/Configuration).

### `{distill}` output

<!-- begin panelset -->

::: {.l-body}
::: {.panelset}

::: {.panel}
### Blank {.panel-name}

Add nothing to end of 4x backticks

`````

````
````

`````

Which returns code formatted like the below:



````
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

:::

<!-- end panel1 -->

::: {.panel}
### `r` {.panel-name}

Add the letter `r` to end of 4x backticks

`````

````r
````

`````

Which returns code formatted like the below:


````r
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

:::

<!-- end panel2 -->

::: {.panel}
### `fortan` {.panel-name}

Add `fortran` to the end of the 4x backticks

`````

````fortran
````

`````

Which returns code formatted like the below:



````fortran
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

:::

<!-- end panel3 -->

::: {.panel}
### `markdown` {.panel-name}

Add `markdown` or `md` to the end of the 4x backticks

`````

````markdown
````

`````

Which returns code formatted like the below:

````markdown
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

:::

<!-- end panel4 -->

::: {.panel}
### `c` {.panel-name}

Add the letter `c` to the end of the 4x backticks

`````

````c
````

`````

Which returns code formatted like the below:


````c
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

:::

<!-- end panel5 -->

::: {.panel}
### `haskell` {.panel-name}

Add `haskell` to the end of the 4x backticks

`````

````haskell
````

`````

Which returns code formatted like the below:


````haskell
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

:::

<!-- end panel6 -->

::: {.panel}
### `python` {.panel-name}

Add `python` to the end of the 4x backticks

`````

````python
````

`````

Which returns code formatted like the below:


````python
```{python}`r ''`
# load library and data
import pandas as pd
import statsmodels.api as sm

mtcars = sm.datasets.get_rdataset("mtcars", "datasets", cache=True).data

# Execute code
(mtcars
  .groupby(['cyl'])
  .agg(['count', 'mean']))
```

````

:::

<!-- end panel7 -->

:::
:::

<!-- end panelset -->


### Xaringan output

```{r, echo = FALSE}
knitr::include_url("https://themockup.blog/posts/2021-08-27-displaying-verbatim-code-chunks-in-xaringan-presentations/xaringan-ex.html")
```


---

## Add more complexity

Now you can also embed essentially entire RMarkdown documents, including inline code. For inline code such as `'r 1 + 1'`, you'll need to use one more technique - also outlined in the [RMarkdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/verbatim-code-chunks.html#show-a-verbatim-inline-expression).

To display inline code like `r knitr::inline_expr('1+1')` you'll need to use `knitr::inline_expr('1+1')` which will convert the raw expression to the expected inline output of `r knitr::inline_expr('1+1')` before it's evaluated by RMarkdown.

A meta example from the RMarkdown Cookbook:

```{r}
code = 'This will show a verbatim inline R expression `r knitr::inline_expr(\'knitr::inline_expr("1+1")\')` in the output.'

cat(code)
```

Which if used as code outputs:

```md
This will show a verbatim inline R expression
`r knitr::inline_expr('knitr::inline_expr("1+1")')` in the output.
```

We can use a full example below.

### Generate inline code

To generate the verbatim inline code we actually used:

```md
We have data about `r knitr::inline_expr('knitr::inline_expr("nrow(penguins)")')` penguins.  Only 
`r knitr::inline_expr('knitr::inline_expr("nrow(penguins) - nrow(smaller)")')` are classified as
`r knitr::inline_expr('knitr::inline_expr("params$species")')`. The distribution of the 
`r knitr::inline_expr('knitr::inline_expr("params$species")')` penguins are shown below:
```

Which generated the code below:

````md

We have data about `r knitr::inline_expr("nrow(penguins)")`  penguins.  Only 
`r knitr::inline_expr("nrow(penguins) - nrow(smaller)")` are classified as
`r knitr::inline_expr("params$species")`. The distribution of the 
`r knitr::inline_expr("params$species")` penguins are shown below:

````

### All together now

Here's YAML + code chunks + inline R expression, all in one code block!


````md
---
title: "Penguins"
date: 2020-08-11
output: html_document
params:
  species: Adelie
---
```{r setup, include = FALSE}`r ''`
library(tidyverse)
library(palmerpenguins)
smaller <- penguins %>% 
  filter(species == params$species, 
         !is.na(body_mass_g))
```
We have data about `r knitr::inline_expr("nrow(penguins)")` penguins.  Only 
`r knitr::inline_expr("nrow(penguins) - nrow(smaller)")` are classified as
`r knitr::inline_expr("params$species")`. The distribution of the 
`r knitr::inline_expr("params$species")` penguins are shown below:

```{r, echo = FALSE}`r ''`
smaller %>% 
  ggplot(aes(body_mass_g)) + 
  geom_histogram(binwidth = 100)
```
````

<aside>

Make sure to keep track of your back ticks as the code can stop looking like _"real"_ code after staring at for a while. I highly recommend checking yourself as you go, because if you miss a backtick it can be quite tricky to "find the needle in the haystack"...

</aside>

Just for fun, I've included a [link to the source code](https://gist.github.com/jthomasmock/771853602d5b15662a36eca237e2113d) for the above embedded RMarkdown so you can explore it.

I also highly recommend exploring the [source code of the RMarkdown Cookbook](https://github.com/yihui/rmarkdown-cookbook/blob/master/05-formatting.Rmd) to figure out how they did all of their inline embedding.

```{r, echo = FALSE, layout = "l-page"}
knitr::include_url("https://bookdown.org/yihui/rmarkdown-cookbook/verbatim-code-chunks.html#", height = "800px")
```

