---
output:
  xaringan::moon_reader:
    nature:
      self_contained: true
      highlightStyle: github
      highlightLines: true
---

```{r,eval = TRUE, echo = FALSE}
xaringanExtra::use_panelset()
```


<!-- begin panelset -->

.panelset[
.panel[.panel-name[Blank]

Add nothing to end of 4x backticks

`````

````
````

`````

Which returns code formatted like the below:



````
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

]

<!-- end panel1 -->

.panel[.panel-name[`r`]

Add the letter `r` to end of 4x backticks

`````

````r
````

`````

Which returns code formatted like the below:


````r
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

]

<!-- end panel2 -->

.panel[.panel-name[`fortran`]

Add `fortran` to the end of the 4x backticks

`````

````fortran
````

`````

Which returns code formatted like the below:



````fortran
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

]

<!-- end panel3 -->

.panel[.panel-name[`markdown`]

Add `markdown` or `md` to the end of the 4x backticks

`````

````markdown
````

`````

Which returns code formatted like the below:

````markdown
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

]

<!-- end panel4 -->

.panel[.panel-name[`c`]

Add the letter `c` to the end of the 4x backticks

`````

````c
````

`````

Which returns code formatted like the below:


````c
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

]

<!-- end panel5 -->

.panel[.panel-name[`haskell`]

Add `haskell` to the end of the 4x backticks

`````

````haskell
````

`````

Which returns code formatted like the below:


````haskell
```{r}`r ''`
# load library
library(dplyr)

# Execute code
mtcars %>% 
  group_by(cyl) %>% 
  summarize(n = n(), mean = mean(mpg)) %>% 
  mutate(cyl_name = paste(cyl, "Cylinders"))
```

````

]

<!-- end panel6 -->

.panel[.panel-name[`python`]

Add `python` to the end of the 4x backticks

`````

````python
````

`````

Which returns code formatted like the below:


````python
```{python}`r ''`
# load library and data
import pandas as pd
import statsmodels.api as sm

mtcars = sm.datasets.get_rdataset("mtcars", "datasets", cache=True).data

# Execute code
(mtcars
  .groupby(['cyl'])
  .agg(['count', 'mean']))
```

````

]

<!-- end panel7 -->

]

<!-- end panelset -->