---
title: "Heatmaps in ggplot2"
description: |
  It's more than just a passing fad
author:
  - name: Thomas Mock
base_url: https://themockup.blog
date: 08-28-2020
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
categories:
  - data visualization
  - ggplot2
  - tidyverse
  - NFL
preview: preview.png
twitter:
  site: "@thomas_mock"
  creator: "@thomas_mock"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggExtra)
library(patchwork)
library(paletteer)
library(scales)
```


Many thanks to [Ethan Douglas](https://twitter.com/ChiefsAnalytics) for sharing his heatmap python code on [OpenSource Football](https://www.opensourcefootball.com/posts/2020-08-22-nfl-pass-location-visualization/)! ^[Douglas (2020, Aug. 21). Open Source Football: NFL Pass Location Visualization. Retrieved from https://mrcaseb.github.io/open-source-football/posts/2020-08-22-nfl-pass-location-visualization/]

This is a similar walkthrough in R + `ggplot2`, but credit for both collecting the data and the original plot were Ethan's!

## Load Everything

```{r, eval = FALSE}
library(tidyverse)
library(ggExtra)
library(patchwork)
library(paletteer)
library(scales)
```


```{r read_data, message=FALSE,warning=FALSE}
souce_url <- "https://raw.githubusercontent.com/ArrowheadAnalytics/next-gen-scrapy-2.0/master/pass_and_game_data.csv"

pass_map_df <- read_csv(souce_url) %>%
  na.omit() %>%
  select(-X1)

glimpse(pass_map_df)
```

## Why not just plot the points?

Why even mess around with `heatmaps` or 2d density plots? Well, we run the risk of overplotting by graphing just the points without thinking or adjusting any of the aesthetics. 

Hadley Wickham in [`ggplot2`: Elegant Graphics for Data Analysis](https://ggplot2-book.org/statistical-summaries.html#overplotting):  

> When the data is large, points will be often plotted on top of each other, obscuring the true relationship. In extreme cases, you will only be able to see the extent of the data, and any conclusions drawn from the graphic will be suspect. This problem is called **overplotting**.

Hadley uses an example with just 2000 data points, which already has a lot of overplotting due to a small sample x-y space. We have about 43,000 passes for this dataset, spread out over a relatively large space. We still need to be careful of overplotting though!

<aside>

Data to Viz has nice article on [Overplotting](https://www.data-to-viz.com/caveat/overplotting.html) as well

</aside>

### Fear the beard

We can display our points with a quick `geom_point()` call but this returns a lot of overplotting albeit in a happy beard shape!

```{r sanity-check, fig.height=4, fig.width=8, layout ='l-body-outset'}
pass_map_df %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_point()
```

We can add 90% transparency with `alpha = 0.1` or set `alpha` as a ratio, and change the color to red, but this still leaves us with difficulty parsing through the high vs medium pass locations. Although the 1/20 ratio gets us a bit farther. The ratio denominator can be though of as the number of points that must stacked to give a solid color.

<aside>

`ggplot2` Book:  

> If you specify alpha as a ratio, the denominator gives the number of points that must be overplotted to give a solid colour. Values smaller than ~ `1/500` are rounded down to zero, giving completely transparent points. ^[https://ggplot2-book.org/statistical-summaries.html#overplotting]

</aside>

Passes around the 5 yard mark appear to be as common as at the 10 and even 20 yard mark in some cases! We know this isn't true from `nflfastR` data. Maybe we can try to get closer views of the actual distributions with `ggMarginal` from the [`ggExtra`](https://github.com/daattali/ggExtra) R package, which provides marginal histograms, box plots, or density plots.

```{r, fig.height=4,fig.width=4}
red_beard <- pass_map_df %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_point(alpha = 0.1, color = "red") +
  labs(title = "Alpha = 0.1 or 90% transparency")

red_beard_ratio <- pass_map_df %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_point(alpha = 1/20, color = "red") +
  labs(title = "Alpha = 1/20, so 20 points must be stacked to give solid color")


red_beard + red_beard_ratio
```

We can add some marginal plots to help out a bit, but that still doesn't solve the problem of honing in on specific areas of interest that well.

```{r marginal-plots, fig.height=4,fig.width=4}
# add density or histogram in the margins
ggMarginal(red_beard, type = "density", fill = "red")
```

```{r marginal-plots2, fig.height=4,fig.width=4}
# add density or histogram in the margins
ggMarginal(red_beard, type = "histogram", fill = "red")
```

### Alternatives to the beard

Since we're actually interested in the density/counts of observations at each x-y coordinate instead of plotting *each* individual point, we also have a few other tools in our toolbox! We can use [`geom_hex()`](https://ggplot2.tidyverse.org/reference/geom_hex.html) or [`geom_bin2d()`](https://ggplot2.tidyverse.org/reference/geom_bin2d.html). Note that there are a few other methods such as `geom_raster()` which require you to provide your own "z" metric, and there are `stat_?` versions of each of these geoms for more custom calculations.

<aside>
If you read the `ggplot2` + overplotting [`ggplot2` book  chapter](https://ggplot2-book.org/statistical-summaries.html#overplotting), you can find more reproducible longer form examples and strategies with `geom_hex()` and `geom_bin2d()`.
</aside>

```{r hex-plot, fig.height=8,fig.width=6}
hex_plot <- pass_map_df %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_hex(
    binwidth = c(1, 1)
  ) +
  scale_fill_gradient(low = "red", high = "yellow") +
  geom_hline(yintercept = c(2, 7), color = "grey") +
  scale_y_continuous(breaks = seq(-10, 60, 5))

hex_plot

bin2d <- pass_map_df %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_bin2d(
    binwidth = c(1, 1)
  ) +
  scale_fill_gradient(low = "red", high = "yellow") +
  geom_hline(yintercept = c(2, 7), color = "grey") +
  scale_y_continuous(breaks = seq(-10, 60, 5))

bin2d
```

We can now see that while there ARE passes all over the field, the majority are around the 2-7 yard mark with small but relatively dense pockets elsewhere. Passes are most frequently thrown at 4, 5 or 6 yards and ~68% of passes are thrown 10 yards or shorter.

```{r not-heatmaps}
# histogram of just Y coord
pass_map_df %>%
  ggplot(aes(x = y_coord)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = c(5)) +
  scale_x_continuous(breaks = seq(-10, 60, 5))

# calc some percentages
pass_by_y <- pass_map_df %>%
  mutate(y_rnd = round(y_coord, digits = 0)) %>%
  count(y_rnd) %>%
  mutate(
    total = sum(n),
    pct_total = n / total,
    roll_total = cumsum(pct_total)
  )

# ~35.6% of passes between 2-7 yards
pass_by_y %>%
  filter(between(y_rnd, 2, 7)) %>%
  summarize(pct = sum(pct_total))

# passes are most commonly thrown at 4-6 yards
# 68% of passes are thrown 10 yards or shorter
pass_by_y %>%
  arrange(desc(pct_total))
```

### Don't bury the lede? 

Another major advantage of `geom_hex` or `geom_bin2d()` is they're remarkably faster for big data than plotting `geom_point()` along with reducing the likelihood of overplotting! For a toy example of about 1.7 million points, `geom_hex()` executes in about 2 sec vs 20 sec with `geom_point()`, and then subsequent 30-60 sec to "draw" the output in the viewer of R/RStudio.

<aside>
Using these techniques are useful, as they are both computationally helpful (aka faster), and can help you better understand plots of "bigger" data.
</aside>

## 2D Density of "smaller" data

So we've covered hex and rectangular 2d bins. To me, these are not as attractive for "small data" like we may see for individual QB plots. Let's take Patrick Mahomes for example, he *only* has ~1,000 passes in this dataset. We can plot with a `geom_hex()` for each 1x1 yard chunk like we did before, but I honestly have trouble determining trends of where he likes to throw with that graphic. 

```{r mahomes-alone, fig.height=8,fig.width=6}
pass_map_df %>%
  filter(str_detect(name, c("Mahomes"))) %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_hex(binwidth = c(1, 1)) +
  scale_y_continuous(breaks = seq(-10, 60, 5))
```

We could always use larger bins, but if we're just trying to see large trends we have another strategy in [`geom_density_2d()`](https://ggplot2.tidyverse.org/reference/geom_density_2d.html).

> Perform a 2D kernel density estimation using [`MASS::kde2d()`](https://rdrr.io/pkg/MASS/man/kde2d.html) and display the results with contours. This can be useful for dealing with overplotting. This is a 2D version of geom_density(). geom_density_2d() draws contour lines, and geom_density_2d_filled() draws filled contour bands.

This essentially fits a polygon around the most frequent points by x/y coordinates, and then colors them according to density. In this case, x/y is basically the position on the field and "z" is the density of how many plots were thrown to that area.


```{r mahomes-2d, fig.height=8,fig.width=6}
pass_map_df %>%
  filter(str_detect(name, c("Mahomes"))) %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_density_2d_filled() +
  scale_y_continuous(breaks = seq(-10, 60, 5))
```

Now if we want to get even more clever, we can use this compare passing heatmaps of specific QBs. We can normalize across the facets, and drop the least frequent passes with specific breaks. I've binned into 10 specific breakpoints, and by setting breaks between 0.1 and 1.0, we also drop the very least frequent passes which otherwise show up as a light fill of the entire remaining plot. Additionally, by adding a horizontal reference line we can pretty clearly see that Carr's most common passes are behind the 5 yard line, while Mahomes has passed beyond the 5 yard line much more frequently.

```{r mahomes-compare, layout="l-body-outset", fig.height=8,fig.width=10}
pass_map_df %>%
  filter(str_detect(name, c("Mahomes|Derek Carr"))) %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_density_2d_filled(
    aes(fill = ..level..),
    contour_var = "ndensity", # normalize to each QBs total passes
    breaks = seq(0.1, 1.0, length.out = 10) # drop the lowest passes
  ) +
  scale_y_continuous(breaks = seq(-10, 60, 5)) +
  facet_wrap(~name) +
  geom_hline(yintercept = 5)
```

This is all well and good, but it's not the prettiest piece of dataviz.

Thanks again to [Ethan Douglas](https://twitter.com/ChiefsAnalytics) and his post on [OpenSourceFootball.com](https://www.opensourcefootball.com/posts/2020-08-22-nfl-pass-location-visualization/) ^[Douglas (2020, Aug. 21). Open Source Football: NFL Pass Location Visualization. Retrieved from https://mrcaseb.github.io/open-source-football/posts/2020-08-22-nfl-pass-location-visualization/], we have a good framework about how to approach building a NFL field as a graph. 

## Build the field

```{r build-field}
not_div_5 <- function(x) {
  # select only elements of the vector not divisible by 5
  x[x %% 5 != 0]
}

center_df <- tibble(
  x_coord = c(rep(-3.1, 60), rep(3.1, 60)),
  y_coord = seq(-14, 59, 1) %>% rep(2) %>% not_div_5(),
  text = "--"
)

# line labels
annotate_df <- tibble(
  x_coord = c(12.88, -12.88) %>% rep(each = 5),
  y_coord = seq(10, 50, 10) %>% rep(2),
  text = seq(10, 50, 10) %>% rep(2) %>% str_replace("(.)(.)", "\\1 \\2"),
  rotation = c(90, 270) %>% rep(each = 5)
)

# yardlines
yardline_df <- tibble(
  y = seq(-15, 60, 5),
  yend = seq(-15, 60, 5),
  x = rep(-56 / 2, 16),
  xend = rep(56 / 2, 16)
)

# sidelines
sideline_df <- tibble(
  y = c(-15.15, -15.15),
  yend = c(60.15, 60.15),
  x = c(-56 / 2, 56 / 2),
  xend = c(-56 / 2, 56 / 2)
)
```

After building some datasets, we can now plot just the field without any other data. This looks nice enough, but I don't want to copy paste it for all the remaining examples, so let's try putting it into a function that we can call with `+ add_field()`!

```{r field_plot, fig.height=8,fig.width=6}
ggplot(data = NULL, aes(x = x_coord, y = y_coord)) +
  coord_cartesian(
    xlim = c(-53.333 / 2, 53.333 / 2),
    ylim = c(-15, 60)
  ) +
  geom_text(
    data = annotate_df, aes(label = text, angle = rotation),
    color = "black", size = 8
  ) +
  geom_segment(
    data = yardline_df, color = "black", size = 1,
    aes(x = x, y = y, xend = xend, yend = yend)
  ) +
  geom_segment(
    x = -56 / 2, y = 0, xend = 56 / 2, yend = 0,
    color = "blue", size = 1, alpha = 0.5
  ) +
  geom_segment(
    data = sideline_df, color = "black", size = 2,
    aes(x = x, y = y, xend = xend, yend = yend)
  ) +
  geom_text(
    data = center_df,
    aes(label = text), color = "black", vjust = 0.32
  ) +
  theme_void()
```

### Wrap the field in a function

We can turn this into a function so that we don't have to copy-paste it all over the place. Note that I'm wrapping it in a list so I can use it with `+` in a sequence of `ggplot2` calls.

I'm also using `front_col` and `back_col` to let us switch from white on black to black on white if needed.


```{r field-function}
add_field <- function() {
  list(
    coord_cartesian(
      xlim = c(-53.333 / 2, 53.333 / 2),
      ylim = c(-15, 60)
    ),
    geom_text(
      data = annotate_df, aes(label = text, angle = rotation),
      color = front_col, size = 8
    ),
    geom_segment(
      data = yardline_df, color = front_col, size = 1,
      aes(x = x, y = y, xend = xend, yend = yend)
    ),
    geom_segment(
      x = -56 / 2, y = 0, xend = 56 / 2, yend = 0,
      color = "blue", size = 1, alpha = 0.5
    ),
    geom_segment(
      data = sideline_df, color = front_col, size = 2,
      aes(x = x, y = y, xend = xend, yend = yend)
    ),
    geom_text(
      data = center_df,
      aes(label = text), color = front_col, vjust = 0.32
    ),
    theme_void(),
    theme(
      strip.text = element_text(size = 20, color = front_col),
      plot.background = element_rect(fill = back_col, color = NA),
      legend.position = "none",
      plot.margin = unit(c(2, 1, 0.5, 1), unit = "cm"),
      plot.caption = element_text(color = front_col),
      plot.title = element_text(color = front_col),
      plot.subtitle = element_text(color = front_col),
      panel.background = element_rect(fill = back_col, color = NA),
      panel.border = element_blank()
    )
  )
}
```

### QB Comparison

First let's filter down to just our two QBs to compare, Patrick Mahomes and Russell Wilson.

```{r create-passer-df, fig.height=8,fig.width=6}
passer_df <- pass_map_df %>%
  filter(str_detect(name, c("Mahomes|Russell"))) %>%
  mutate(name = factor(name, levels = c("Patrick Mahomes", "Russell Wilson"))) %>%
  select(name, x_coord, y_coord)

passer_df %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_density2d_filled() +
  theme(legend.position = "none")
```

We'll specify fill and color to both scale with the level/density and normal density peaks across our plots, and finally set our breaks to drop the lowest bin of passes.

```{r initial-map, fig.height=10,fig.width=10}
pass_map <- passer_df %>%
  ggplot(aes(x = x_coord, y = y_coord)) +
  geom_density_2d_filled(
    aes(fill = ..level.., color = ..level..),
    contour_var = "ndensity", # normalize across facets
    breaks = seq(0.1, 1.0, length.out = 10)
  ) +
  facet_wrap(~name)

pass_map
```

We can quickly add the field background to this with our function `add_field()`!

```{r add-field-plot, fig.height=10,fig.width=10, layout='l-body-outset'}
back_col <- "white"
front_col <- "black"

pass_map +
  add_field()
```

### Specify Color schemes

While that's essentially our final graph perhaps you don't want to use `viridis` which is the default color scheme. We can generate custom color palettes or use a pre-built color palette via the [`paleteer`](https://github.com/EmilHvitfeldt/paletteer) R package. Note that the 3 color palettes I create all do essentially the same thing but it's:  

* Building your own custom color sequence with `grDevices::colorRampPalette()`  
* Returning a pre-built palette w/ `paletteer::paletteer_d()`  
* Expanding a pre-built palette to be longer with `colorRampPalette`

```{r colors}
heat_colors <- grDevices::colorRampPalette(c("#800026FF", "#FC4E2AFF", "#FEB24CFF", "#FFFFCCFF"))(10)

heat_palette <- paletteer::paletteer_d("RColorBrewer::YlOrRd", n = 9, direction = -1)

heat_colors_interpolated <- colorRampPalette(paletteer::paletteer_d("RColorBrewer::YlOrRd", n = 9, direction = -1))(10)

heat_colors %>% scales::show_col()
heat_palette %>% scales::show_col()
heat_colors_interpolated %>% scales::show_col()
```

```{r use-colors, fig.height=12,fig.width=12, layout='l-body-outset'}
pass_map +
  add_field() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill", "color"))
```

#### And also in black!

```{r use-colors2, fig.height=12,fig.width=12, layout='l-body-outset'}
back_col <- "black"
front_col <- "white"

pass_map +
  add_field() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill", "color"))
```


## PFF_Moo style Field

[PFF_Moo](https://twitter.com/PFF_Moo/status/1235681100837486592?s=20) takes a different approach to field lines, which can be recreated below. ^[https://twitter.com/PFF_Moo/status/1235681100837486592?s=20]

```{r field_plot_moo, fig.height=10,fig.width=6}
not_div_5 <- function(x) {
  # select only elements of the vector not divisible by 5
  x[x %% 5 != 0]
}

center_df <- tibble(
  x_coord = c(rep(-3.1, 60), rep(3.1, 60)),
  y_coord = seq(-14, 59, 1) %>% rep(2) %>% not_div_5(),
  text = "--"
)

# line labels
horiz_yd_df <- tibble(
  x_coord = c(12.88, -12.88) %>% rep(each = 14),
  y_coord = seq(-10, 55, 5) %>% rep(2),
  text = seq(-10, 55, 5) %>% rep(2)
)

# yardlines
yardline_df <- tibble(
  y = seq(-15, 60, 5),
  yend = seq(-15, 60, 5),
  x = rep(-56 / 2, 16),
  xend = rep(56 / 2, 16)
)

# sidelines
sideline_df <- tibble(
  y = c(-15.15, -15.15),
  yend = c(60.15, 60.15),
  x = c(-56 / 2, 56 / 2),
  xend = c(-56 / 2, 56 / 2)
)

add_moo_field <- function() {
  list(
    coord_cartesian(
      xlim = c(-53.333 / 2, 53.333 / 2),
      ylim = c(-15, 60)
    ),
    geom_segment(
      data = yardline_df, color = front_col, size = 0.5,
      linetype = "dashed", alpha = 0.5,
      aes(x = x, y = y, xend = xend, yend = yend)
    ),
    geom_segment(
      aes(x = -56 / 2, y = 0, xend = 56 / 2, yend = 0),
      color = "blue", size = 1
    ),
    geom_segment(
      data = sideline_df, color = front_col, size = 4,
      aes(x = x, y = y, xend = xend, yend = yend)
    ),
    geom_text(
      data = center_df,
      aes(label = text), color = front_col, vjust = 0.32
    ),
    geom_text(
      data = horiz_yd_df, aes(label = text),
      color = front_col, size = 4, fontface = "bold"
    ),
    theme_void(),
    theme(
      strip.text = element_text(size = 20, color = front_col),
      plot.background = element_rect(fill = back_col, color = NA),
      legend.position = "none",
      plot.margin = unit(c(2, 1, 0.5, 1), unit = "cm"),
      plot.caption = element_text(color = front_col),
      plot.title = element_text(color = front_col),
      plot.subtitle = element_text(color = front_col),
      panel.background = element_rect(fill = back_col, color = NA),
      panel.border = element_blank()
    )
  )
}

back_col <- "white"
front_col <- "black"


ggplot(pass_map_df, aes(x = x_coord, y = y_coord)) +
  geom_density_2d_filled(
    aes(fill = ..level..),
    contour_var = "ndensity", # normalize to each QBs total passes
    breaks = seq(0.1, 1.0, length.out = 10) # drop the lowest passes
  ) +
  add_moo_field()
```

## Get the raw density estimates

Lastly, we can also get the raw density measures to use against the points for example or for other more computational measures as opposed to just graphics. ^[Function adapted from [Kamil Slowikowski](https://slowkow.com/notes/ggplot2-color-by-density/)] Note that this is for ALL the QBs at once.


```{r dens-calc,fig.height=6, fig.width=6}

get_density <- function(x, y, ...) {
  density_out <- MASS::kde2d(x, y, ...)
  int_x <- findInterval(x, density_out$x)
  int_y <- findInterval(y, density_out$y)
  comb_int <- cbind(int_x, int_y)
  return(density_out$z[comb_int])
}

density_map <- pass_map_df %>% 
  select(x_coord, y_coord, name) %>% 
  # function works in dplyr!
  mutate(density = get_density(x_coord, y_coord, n = 100))

density_map %>% 
  ggplot(aes(x = x_coord, y = y_coord, color = density)) +
  geom_point(alpha = 0.2) +
  scale_color_gradient(low = "red", high = "yellow")
```
  
<aside>

Here's the hexplot for comparison:

```{r, echo = FALSE, fig.width=5, fig.height=6}
hex_plot
```

</aside>

### Compare 2D Density

Now for this, we're actually creating two separate 2 dimensional density estimates, and then subtracting the density of QB2's passes from QB1's.

This is requires us to play around a bit more with matrices, which is a good exercise! I'm going to keep working to see if this could fit into a `tibble` oriented workflow, but so far it's a matrix up to the end. Credit to eipi10 for code I adapted from [Stack Overflow](https://stackoverflow.com/questions/28521145/r-calculate-and-plot-difference-between-two-density-countours). ^[https://stackoverflow.com/questions/28521145/r-calculate-and-plot-difference-between-two-density-countours]

```{r compare_qb_density}
qb_density_compare <- function(pass_df, qb1_name, qb2_name, n = 200){
  
  # filter to qb1
  qb1 <- pass_df %>% 
    select(x_coord, y_coord, name) %>% 
    filter(str_detect(name, qb1_name))
  
  #filter to qb2
  qb2 <- pass_df %>% 
    select(x_coord, y_coord, name) %>% 
    filter(str_detect(name, qb2_name))
  
  # get x/y coords as vectors
  qb1_x <- pull(qb1, x_coord)
  qb1_y <- pull(qb1, y_coord)
  
  # get x/y coords as vectors
  qb2_x <- pull(qb2, x_coord)
  qb2_y <- pull(qb2, y_coord)

  # get x and y range to compute comparisons across
  x_rng = range(c(qb1_x, qb2_x))
  y_rng = range(c(qb1_y, qb2_y))
  
  # Calculate the 2d density estimate over the common range
  d2_qb1 = MASS::kde2d(qb1_x, qb1_y, lims=c(x_rng, y_rng), n=n)
  d2_qb2 = MASS::kde2d(qb2_x, qb2_y, lims=c(x_rng, y_rng), n=n)
  
  # create diff df
  qb_diff <- d2_qb1
  
  # matrix subtraction density from qb2 from qb1
  qb_diff$z <- d2_qb1$z - d2_qb2$z
  
  # add matrix col names
  colnames(qb_diff$z) = qb_diff$y
  
  #### return tidy tibble ####
  qb_diff$z %>% 
    # each col_name is actually the y_coord from the matrix
    as_tibble() %>% 
    # add back the x_coord
    mutate(x_coord= qb_diff$x) %>% 
    pivot_longer(-x_coord, names_to = "y_coord", values_to = "z") %>% 
    mutate(y_coord = as.double(y_coord))

}
```

### Call the function

Now we can just call the function to generate a beautiful dataset to plot or work with!

Let's compare... Mahomes vs Rodgers.

```{r compare_qb_plot, fig.height = 10, fig.width = 8, layout='l-body-outset'}
compared_z <- qb_density_compare(pass_map_df, "Mahomes", "Rodgers", n = 200) 

compared_z %>% 
  ggplot(aes(x_coord, y_coord)) +
  geom_raster(aes(x_coord, y_coord, fill=z))  +
  stat_contour(aes(colour=..level.., z = z), 
               breaks = seq(min(compared_z$z), max(compared_z$z), length.out = 10)
               ) +
  scale_fill_gradient2(low="blue",mid="white", high="red", midpoint=0) +
  scale_colour_gradient2(low=muted("blue"), mid="white", high=muted("red"), midpoint=0) +
  add_moo_field() +
  theme(legend.position = "bottom", legend.key.width = unit(2, "cm"),
        plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(size = 12, hjust = 0.5),
        plot.caption = element_text(face = "bold")) +
  labs(title = "Mahomes (QB1) vs Rodgers (QB2)",
       subtitle = "Red = More by QB1, Blue = More by QB2",
       caption = "\n\nPlot: @thomas_mock | Data: @ChiefsAnalytics") +
  guides(colour=FALSE) 
```


Try it out for yourself!