---
title: "Embedding custom HTML in gt tables"
description: |
  HTML is basically a superpower.
author:
  - name: Thomas Mock
    url: https://twitter.com/thomas_mock
base_url: https://themockup.blog
date: 10-31-2020
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
categories:
  - data visualization
  - tables
  - tidyverse
preview: preview.png
twitter:
  site: "@thomas_mock"
  creator: "@thomas_mock"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(formattable)
library(kableExtra)
library(gt)
library(sparkline)
sparkline(0)
```

### `gt` loves HTML

`gt` really does love HTML, and for now, `gt` is HTML-first which is great since HTML can do SOO many things in R Markdown, `shiny`, and elsewhere.

A `gt` table can always be saved as an `.png` or `.pdf` file if you need to incorporate it in non-HTML content. Lastly, `gt` does have preliminary support for RTF, and could export to a PDF via `gtsave()`.

`gt` allows for HTML to be incorporated in MANY different locations with `gt::html()`. For example, we can use a `<span style>` to color the title for this table, and another `<span style>` to change the font size and weight for one off column. This really just scratches the surface of what is possible, and mostly you are free to use a lot of your creativity in building these tables.

As a note - we are somewhat playing with fire as we're doing quite a bit of nesting for HTML. If you run into problems getting some of these tables to build in certain environments,you can try a `R Markdown` chunk with `results='asis'`, along with `gt() %>% as_raw_html(inline_css = TRUE)`.

Also, if you want to learn a bit more CSS/HTML to adjust tables beyond what I've shown below - check out either [Mozilla Docs](https://developer.mozilla.org/en-US/docs/Web/CSS) or the [W3schools](https://www.w3schools.com/).

```{r}
nfl_qbr <- espnscrapeR::get_nfl_qbr(2020) %>% 
  slice(1:10)

ex_tab <- nfl_qbr %>% 
  select(rank, last_name, team, qbr_total, qb_plays, pass, run) %>% 
  gt() %>% 
  tab_header(
    title = gt::html("<span style='color:red'>ESPN's QBR for 2020</span>")
  ) %>% 
  cols_label(
    qbr_total = gt::html(
      "<span style ='font-weight:bold;font-size:20px'>QBR</span>")
  )
ex_tab
```

### More HTML!

So that's cool to see where things can be changed, but let's walk through a bit more engaging example. Here we're going to merge some columns for the Player's last name + team. 

```{r}
ex_tab <- nfl_qbr %>% 
  select(rank, last_name, team, qbr_total, qb_plays, pass, run) %>% 
  gt() 

ex_tab %>%
  cols_merge(
    columns = vars(last_name, team)
  )
```

This saves us some space since we're dropping a column, but isn't the prettiest thing. Let's use an anonymous function and `text_transform` to change the styling of our player's name/team with `<span style>` along with small caps, different font colors and sizes.

```{r}
ex_tab %>%
  cols_merge(
    columns = vars(last_name, team)
  ) %>% 
  text_transform(
    locations = cells_body(
      columns = vars(last_name)
    ),
    fn = function(x){
      name <- word(x, 1)
      team <- word(x, -1)
      glue::glue(
        "<div><span style='font-weight:bold;font-variant:small-caps;font-size:14px'>{name}</div>
        <div><span style ='font-weight:bold;color:grey;font-size:10px'>{team}</span></div>"
      )
    }
  )
```

This is starting to look better! However, since we stacked it, the rows are very tall, a bit too tall in my opinion. We can use `line-height` inside the `<div>` now to decrease the vertical space between our words.

```{r}
ex_tab %>%
  cols_merge(
    columns = vars(last_name, team)
  ) %>% 
  text_transform(
    locations = cells_body(
      columns = vars(last_name)
    ),
    fn = function(x){
      name <- word(x, 1)
      team <- word(x, -1)
      glue::glue(
        "<div style='line-height:10px'><span style='font-weight:bold;font-variant:small-caps;font-size:14px'>{name}</div>
        <div style='line-height:12px'><span style ='font-weight:bold;color:grey;font-size:10px'>{team}</span></div>"
      )
    }
  ) %>% 
  tab_options(
    data_row.padding = px(5),
  )
```

While we did that ALL within `gt`, we could also have made similar changes by writing some HTML with functions inside `mutate` ahead of sending it to `gt`!

```{r}
# function to incorporate player name + team
combine_word <- function(name, team){
      glue::glue(
        "<div style='line-height:10px'><span style='font-weight:bold;font-variant:small-caps;font-size:14px'>{name}</div>
        <div style='line-height:12px'><span style ='font-weight:bold;color:grey;font-size:10px'>{team}</span></div>"
      )
    }

nfl_qbr %>% 
  select(rank, short_name, team, qbr_total, qb_plays, pass, run) %>% 
  mutate(
    combo = combine_word(short_name, team),
    combo = map(combo, gt::html)
    ) %>% 
  select(rank, combo, everything(), -short_name, -team) %>% 
  gt() %>% 
  cols_align(
    align = "left",
    columns = vars(combo)
  ) %>% 
  tab_options(
    data_row.padding = px(5)
  )
```

So that's really cool and allows you to do some creative things with HTML-based content. What else can we do with HTML?

### `kableExtra` integration

The fantastic [`kableExtra` package](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Overview) has some sparkline-esque graphing capabilities that export as SVG, meaning they can be integrated into HTML.

Note that while I love `gt`, `kableExtra` is again a great package in it's own right and has more mature LaTeX integration today. If you REALLY have to use PDF/LaTex, it's the better choice over `gt` today.

`kableExtra` approaches inline plots with the `spec_plot()` family of functions.

### `kableExtra` example

Here's a quick example from `kableExtra`, which can be adapted to work in `gt`, mainly incorporating an inline boxplot into the table.

```{r, message=FALSE, warning=FALSE}
library(kableExtra)
library(gt)
library(tidyverse)
```

`kableExtra` method, adapted from the [great guide by Hao Zhu](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Insert_Images_into_Columns). 

```{r}
# first split the data by cylinders
mpg_list <- split(mtcars$mpg, mtcars$cyl)

mpg_list

# pipe the 
data.frame(
  cyl = c(4,6,8),
  boxplot = ""
  ) %>% 
  kbl(booktabs = TRUE) %>%
  kable_paper(full_width = FALSE) %>%
  column_spec(2, image = spec_boxplot(mpg_list, width = 300, height = 70))
```

### `gt` + `kableExtra`

We can adapt a similar idea for `gt`, here we are using `mutate` calls ahead of time to prep the data. Here we are going to keep all the data in a pipe, rather than having to `split` it and reference a dataset external to our table. We can essentially nest the same `mpg` column by group, keeping it in a single `tibble` this time.

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarize(data = list(mpg), .groups = "drop")
```

Then we can create a range to set baselines for MPG, and then use `kableExtra::spec_plot()` to embed an inline sparkline. Note we have to use `purrr::map()` here to apply the function iteratively across each row.

```{r}
mpg_rng <- range(mtcars$mpg)

mtcars %>% 
  group_by(cyl) %>% 
  summarize(data = list(mpg), .groups = "drop") %>% 
  mutate(
    plot = map(data, ~spec_plot(.x, ylim = mpg_rng, same_lim = TRUE, width = 300, height = 70)),
    plot = map(plot, "svg_text"),
    plot = map(plot, gt::html)
    ) %>% 
  select(-data) %>% 
  gt()
```

Now that I've showed that it's possible, what are we actually doing? `kableExtra::spec_plot()` creates a plot in base R, and then returns it as either svg or pdf, which means it can be compatible with either HTML or LaTeX. Remember the `mpg_list` we created by splitting the `mpg` column into a list of vectors by `cyl`? 

```{r}
mpg_list %>% str()
```

We can create plots for each `cyl` and then pull the `cyl == 4` plot and look at it's structure.

```{r}
spec_plot(mpg_list) %>% 
  pluck("4") %>% 
  str()
```

We see that it returns a list object, with mostly metadata about what parameters were passed to the function. The part we *really* want is the `svg_text` since that has the `xml` code to generate our inline plot. We can pull out the `svg_text` list item from our list of lists by calling `map("svg_text")`. Now we can see each of the svg-plots, one for each cylinder group!

```{r}
spec_plot(mpg_list) %>% 
  map("svg_text") %>% 
  str()
```

So now that we have the specific item of interest we need to let `gt` "know" to treat this as HTML and not just a random character string. We can call `map()` one more time and apply the `gt::html()` function to each svg plot.

```{r}
spec_plot(mpg_list) %>% 
  map("svg_text") %>% 
  map(gt::html) %>% 
  str()
```

And that's why the following code works:  
- Group by cylinder  
- Summarize down to a list-column of the respective MPG column by cylinder  
- Create the `spec_plot` object  
- Extract the `svg_text`  
- Recognize the `svg_text` as HTML  
- `gt` takes the HTML and parses it  

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarize(data = list(mpg), .groups = "drop") %>% 
  mutate(
    plot = map(data, ~spec_plot(.x, ylim = mpg_rng, same_lim = TRUE, width = 300, height = 70)),
    plot = map(plot, "svg_text"),
    plot = map(plot, gt::html)
    ) %>% 
  select(-data) %>% 
  gt()
```

### Do it all in `gt`

While that works just fine, it assumes that you create the content ahead of time, before incorporating it into `gt`. However, you can also approach it from within `gt` itself.

For the next one, I have a more general function to use.

The custom function `gt_plot()`:  
- Takes the table data from `gt`  
- You specify a specific column  
- You specify external data to plot  
- Specify what type of plot  
- Optionally pass additional arguments to `spec_plot` with `...`  

```{r}
gt_plot <- function(table_data, column, plot_data, plot_fun, ...){
  text_transform(
    table_data,
    # note the use of {{}} here - this is tidy eval
    # that allows you to indicate specific columns
    locations = cells_body(columns = vars({{column}})),
    fn = function(x){
      plot <- map(plot_data, plot_fun, width = 300, height = 70, same_lim = TRUE, ...)
      plot_svg <- map(plot, "svg_text")
      map(plot_svg, gt::html)
    }
  )
}
```

Note that again, my table "data" is pretty minimal, and I've got the data externally as our `mpg_list` object we created earlier.

```{r}
mpg_list %>% str()
```

```{r}
tibble(cyl = c(4,6,8), boxplot = "") %>% 
  gt() %>% 
  gt_plot(
    column = boxplot,  # column to create plot in 
    plot_data = mpg_list, # external data to reference
    plot_fun = spec_boxplot,  # which plot fun
    lim = mpg_rng # range applied
    )
```

We can quickly switch from a boxplot to a sparkline, just by changing the `plot_fun` argument to `spec_plot`. Also since I passed ellipses (`...`) to the `spec_plot()` function we can also use some additional arguments to change the line-color to black, and make the max/min points to be a bit larger.

```{r}
tibble(cyl = c(4,6,8), boxplot = "") %>% 
  gt() %>% 
  gt_plot(
    column = boxplot,  # column to create plot in 
    plot_data = mpg_list, # external data to reference
    plot_fun = spec_plot,  # which plot fun
    ylim = mpg_rng, # range applied,
    col = "black", # change color of line
    cex = 5 # change size of points
    )
```

This works with all the `kableExtra` inline plot functions! Note that we are just varying the `ylim` on the line/points 1 vs 2, where the mpg_line1/mpg_points1 share a common y-axis, and line2/points2 have their own y-axis.

```{r, layout = "l-page"}
tibble(
  cyl = c(4,6,8), 
  boxplot = "", mpg_hist = "", mpg_line1 = "", 
  mpg_line2 = "", mpg_points1 = "", 
  mpg_points2 = "", mpg_poly = ""
  ) %>% 
  gt() %>% 
  gt_plot(column = boxplot, plot_data = mpg_list, plot_fun = spec_boxplot, lim = mpg_rng) %>% 
  gt_plot(column = mpg_hist, plot_data = mpg_list, plot_fun = spec_hist, lim = mpg_rng) %>% 
  gt_plot(column = mpg_line1, plot_data = mpg_list, plot_fun = spec_plot, ylim = mpg_rng) %>% 
  gt_plot(column = mpg_line2, plot_data = mpg_list, plot_fun = spec_plot) %>% 
  gt_plot(column = mpg_points1, plot_data = mpg_list, plot_fun = spec_plot, type = "p", ylim = mpg_rng, cex = 4) %>% 
  gt_plot(column = mpg_points2, plot_data = mpg_list, plot_fun = spec_plot, type = "p", cex = 4) %>% 
  gt_plot(column = mpg_poly, plot_data = mpg_list, plot_fun = spec_plot, polymin = 5, ylim = mpg_rng)
```

### Use a single source of data

OK so we now have a function, but we're referencing an external data object, rather than data within the "table" itself - not ideal!

Can we use our `group_by` + `summarize` as list from before? (Spoiler = `nope`)

```{r}
# doesn't work
mtcars %>% 
  group_by(cyl) %>% 
  summarize(data = list(mpg), .groups = "drop") %>% 
  gt() %>% 
  text_transform(
    locations = cells_body(columns = vars(data)),
    fn = function(x){
    plot = map(x, ~spec_plot(.x, ylim = mpg_rng, same_lim = TRUE, width = 300, height = 70))
    plot = map(plot, "svg_text")
    plot = map(plot, gt::html)
    }
  ) 
```

Nope - but it does give us a decent error message!

```
1: In xy.coords(x, y, xlabel, ylabel, log) : NAs introduced by coercion
2: In which.min(y) : NAs introduced by coercion
3: In which.max(y) : NAs introduced by coercion
```

There seems to be a type conversion - NAs are being returned where we expect numeric data to create the x-y coordinates for the plot. Let's dive a bit closer into what happens when we call `text_transform()`. I'm calling `str()` inside our `text_transform()` now to expose what the data itself looks like.

```{r, eval = FALSE}
# doesn't work
mtcars %>% 
  group_by(cyl) %>% 
  summarize(mpg_data = list(mpg), .groups = "drop") %>% 
  gt() %>% 
  text_transform(
    locations = cells_body(columns = vars(mpg_data)),
    fn = function(x){
      str(x)
    })
```

```
List of 3
 $ : chr "22.8, 24.4, 22.8, 32.4, 30.4, 33.9, 21.5, 27.3, 26.0, 30.4, 21.4"
 $ : chr "21.0, 21.0, 21.4, 18.1, 19.2, 17.8, 19.7"
 $ : chr "18.7, 14.3, 16.4, 17.3, 15.2, 10.4, 10.4, 14.7, 15.5, 15.2, 13.3, 19.2, 15.8, 15.0"
Error: Assigned data `*vtmp*` must be compatible with existing data.
x Existing data has 3 rows.
x Assigned data has 0 rows.
ℹ Only vectors of size 1 are recycled.
```

This tells us something interesting! It's combined all the vectors into a character string separated by commas. No wonder our graph can't understand its xy coords, it is passed as one long text string!

Now if we're tricky, we can get at the `guts` of `gt` since it's just a list object. There's quite a bit there inside the `gt` object, but the first list item is arguably the most important! We have the raw data as `_data`!

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarize(mpg_data = list(mpg), .groups = "drop") %>% 
  gt() %>% 
  str(max.level = 1)
```

We can `pluck()` the raw underlying data itself from `gt`, extract the `mpg_data` column, and could work with it in our function.

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarize(mpg_data = list(as.double(mpg)), .groups = "drop") %>% 
  gt() %>% 
  pluck("_data", "mpg_data") %>% 
  str()
```

So let's try it out! Remember, we're using `pluck()` to get the dataframe from `gt`'s list object, and then pulling out the `mpg_data` column from it.

```{r}
# works now
mtcars %>% 
  group_by(cyl) %>% 
  summarize(mpg_data = list(mpg), .groups = "drop") %>% 
  gt() %>% 
  text_transform(
    locations = cells_body(columns = vars(mpg_data)),
    fn = function(x){
      data_in = pluck(., "_data", "mpg_data")
      plot = map(data_in, ~spec_plot(.x, ylim = mpg_rng, same_lim = TRUE, width = 300, height = 70))
      plot = map_chr(plot, "svg_text")
    })
```

That worked beautifully!

An alternative that doesn't require going into the `gt` object itself and coerces the character string back into numeric.

```{r}
# WORKS
mtcars %>%
  group_by(cyl) %>%
  summarize(mpg_data = list(as.double(mpg)), .groups = "drop") %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = vars(mpg_data)),
    fn = function(x) {
      # split the strings at each comma
      split_data <- str_split(x, ", ")
      # convert to type double
      data <- map(split_data, as.double)
      # create the plot
      plot <- map(data, ~ spec_plot(.x, ylim = mpg_rng, same_lim = TRUE, width = 300, height = 70))
      # extract the svg item
      map(plot, "svg_text")
    }
  )
```

Ok so we've shown that it's possible to do that either way, so let's rewrite our function!

```{r}
gt_plot <- function(table_data, plot_col, data_col, plot_fun, ...){
  # save the data extract ahead of time 
  # to be used in our anonymous function below
  data_in = pluck(table_data, "_data", data_col)
  
  text_transform(
    table_data,
    # note the use of {{}} here - this is tidy eval
    # that allows you to indicate specific columns
    locations = cells_body(columns = vars({{plot_col}})),
    fn = function(x){
      plot <- map(data_in, plot_fun, width = 300, height = 70, same_lim = FALSE, ...)
      plot_svg <- map(plot, "svg_text")
      map(plot_svg, gt::html)
    }
  )
}
```

This function will now work exactly as expected with the grouped list data columns!

```{r}
# works!
mtcars %>% 
  group_by(cyl) %>% 
  summarize(mpg_data = list(mpg), .groups = "drop") %>% 
  gt() %>% 
  # note you can leave mpg_data unquoted for the tidyeval
  # but have to quote mpg_data for the pluck
  gt_plot(mpg_data, "mpg_data", plot_fun = spec_plot)
```

### Interactive `sparklines`

So the embedded sparklines as shown above are fantastic, quick and robust, but they're static. Since we're focusing on HTML content, why don't we also see if we can get javascript enabled interactivity?

Quick example of this working below, but note you need to call `sparkline(0)` somewhere ahead of time in your RMarkdown doc to load the javascript library dependency. Also, if you try to view this interactively it will look like it failed and didn't pass anything through, but it will work when the RMarkdown is knit.

```{r}
tibble(
  var = c("mpg", "wt"),
  sparkline1 = "",
  sparkline2 = "",
  box = ""
) %>% 
  gt() %>% 
  text_transform(
    locations = cells_body(vars(sparkline1)),
    fn = function(x){
      sparkline <- map(list(mtcars$mpg, mtcars$wt), ~spk_chr(values = .x, chartRangeMin = 0))
      map(sparkline, gt::html)
    }
  ) %>% 
  text_transform(
    locations = cells_body(vars(sparkline2)),
    fn = function(x){
      sparkline <- map(list(mtcars$mpg, mtcars$wt), ~spk_chr(values = .x, type = "bar", chartRangeMin = 0))
      map(sparkline, gt::html)
    }
  ) %>% 
  text_transform(
    locations = cells_body(vars(box)),
    fn = function(x){
      sparkline <- map(list(mtcars$mpg, mtcars$wt), ~spk_chr(values = .x, type = "box", chartRangeMin = 0))
      map(sparkline, gt::html)
    }
  )
```

While we're likely to only be using this in a table once per each , I do want to try and create a function so that we don't have to re-write these each time and could potentially roll it into a package.

```{r}
gt_spark <- function(table_data, plot_col, data_col){
  # save the data extract ahead of time 
  # to be used in our anonymous function below
  data_in = pluck(table_data, "_data", data_col)
  
  text_transform(
    table_data,
    # note the use of {{}} here - this is tidy eval
    # that allows you to indicate specific columns
    locations = cells_body(columns = vars({{plot_col}})),
    fn = function(x){
      sparkline_plot <- map(
        data_in, 
        ~spk_chr(values = .x, chartRangeMin = 0)
        )
      
      map(sparkline_plot, gt::html)
    }
  )
}
```

We can then apply the function to work very succinctly, referencing only the internal list-column data.

```{r}
# works!
mtcars %>% 
  group_by(cyl) %>% 
  summarize(mpg_data = list(mpg), .groups = "drop") %>% 
  gt() %>% 
  # note you can leave mpg_data unquoted for the tidyeval
  # but have to quote mpg_data for the pluck
  gt_spark(mpg_data, "mpg_data")
```

### Forest

You can also make forest-plot like tables in `gt`, note that this code is taken essentially verbatim from [`kableExtra`'s documentation](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Insert_Images_into_Columns), just adapted to work in `gt`.

```{r}
coef_table <- data.frame(
  Variables = c("var 1", "var 2", "var 3"),
  Coefficients = c(1.6, 0.2, -2.0),
  Conf.Lower = c(1.3, -0.4, -2.5),
  Conf.Higher = c(1.9, 0.6, -1.4)
) 

tibble(
  Variable = coef_table$Variables,
) %>%
  mutate(
    image = spec_pointrange(
      x = coef_table$Coefficients, 
      xmin = coef_table$Conf.Lower, 
      xmax = coef_table$Conf.Higher, 
      vline = 0,
      width = 250,
      cex = .75,
      col = "black",
      pch = 16
      )
    ) %>% 
  mutate(
    image = map(image, "svg_text"),
    image = map(image, ~gt::html(as.character(.x)))
  ) %>% 
  gt()
```

We can show a bit more robust example from a recent question by [Silvia Canelón
- @spcanelon](https://twitter.com/spcanelon/status/1320832482334396416?s=21)

<details>
<summary><strong>Code to generate fake data </summary>  

```{r}
coef_table <- tibble(
  group = c(
    "",
    rep("Sex", 2),
    rep("Age", 4),
    rep("Body-Mass index", 2),
    rep("Race", 3),
    rep("Baseline statin treatment", 2),
    rep("Intensity of statin treatment", 2),
    rep("Metabolic disease", 3),
    rep("Renal function", 3)
  ),
  subgroup = c(
    "All Patients",
    "Male", "Female",
    "<65 yr", ">= 65 yr", "<75 yr", ">=75 yr",
    "<=Median", ">Median",
    "White", "Black", "Other",
    "Yes", "No",
    "High", "Not high",
    "Diabetes", "Metabolic syndrome", "Neither",
    "Normal", "Mild impairment", "Moderate impairment"
  ),
  Inclisiran = c(
    781, 535,246,297,484,638,143,394,387,653,110,18,701,80,538,243,371,195,215,395,269,113
  ),
  Placebo = c(
    780,548,232,333,447,649,131,385,394,685,87,8,692,88,546,234,331,207,242,410,260,107
  ),
  coefficients = c(-60,-55,-68,-58,-55,-57,-58,-55,-48,-58,-57,-49,-44,-58,-55,-57,-54,-52,-54,-53, -54,-52)
  ) %>% 
  mutate(
    conf_range = runif(22, min = 5, max = 10),
    conf_lower = coefficients - conf_range,
    conf_higher = coefficients + conf_range
  ) %>%
  mutate(
    image = spec_pointrange(
      x = coefficients, 
      xmin = conf_lower, 
      xmax = conf_higher, 
      same_lim = TRUE,
      lim = c(-100, 25),
      vline = 0,
      width = 550,
      cex = .75,
      col = "black"
      )
    )
```

</details>

Here's the code to create a quick table with a zero-indicated line, and some randomgly generated 

```{r}
coef_table %>% 
  select(-coefficients, -contains("conf")) %>% 
  mutate(
    image = map(image, "svg_text"),
    image = map(image, ~gt::html(as.character(.x)))
  ) %>% 
  select(group:Placebo, pct_diff = image) %>% 
  gt(
    groupname_col = "group",
    rowname_col = "subgroup"
  ) %>% 
  opt_row_striping() %>% 
  tab_options(
    data_row.padding = px(3)
  )
```

### More custom HTML work

For the next section, I'll be showing some functions that are mostly adapted from Greg Lin's fantastic examples for the [`reactable` package Cookbook](https://glin.github.io/reactable/articles/cookbook/cookbook.html). I love `reactable`, but want to show how some of the same ideas can translate in to mostly static tables as well. Note that some of the `tags$`, `div`, etc are from the `htmltools` package, and you can generally write your own HTML by hand if you wanted.



<details>
<summary><strong>Function to add tooltip to a table column label</summary>  

```{r}
library(htmltools)

# Add tooltip to column labels
with_tooltip <- function(value, tooltip) {
  tags$abbr(style = "text-decoration: underline; text-decoration-style: solid; cursor: question; color: blue",
            title = tooltip, value)
}
```

</details>

<br>

<details>
<summary><strong>Function that creates a star rating scale from 0-5</summary>  

```{r}
# note you could use ANY font-awesome logo
# https://fontawesome.com/cheatsheet
rating_stars <- function(rating, max_rating = 5) {
  rounded_rating <- floor(rating + 0.5)  # always round up
  stars <- lapply(seq_len(max_rating), function(i) {
    if (i <= rounded_rating) fontawesome::fa("star", fill= "orange") else fontawesome::fa("star", fill= "grey")
  })
  label <- sprintf("%s out of %s", rating, max_rating)
  div_out <- div(title = label, "aria-label" = label, role = "img", stars)
  
  as.character(div_out) %>% 
    gt::html()
}
```

</details>

<br>

<details>
<summary><strong>`fontawesome` package for inline icons</summary>  

```{r}
rank_chg <- function(change_dir){
  if (change_dir == "increase") {
    logo_out <- fontawesome::fa("arrow-up", fill = "blue")
  } else if (change_dir == "decrease"){
    logo_out <- fontawesome::fa("arrow-down", fill = "red")
  }
  
  logo_out %>% 
    as.character() %>% 
    gt::html()
  
}
```

</details>

<br>

<details>
<summary><strong>Create a "badge" style label with a specific color, and round edges.</summary>  

```{r}
add_cyl_color <- function(cyl){
      add_color <- if (cyl == 4) {
        "background: hsl(116, 60%, 90%); color: hsl(116, 30%, 25%);"
      } else if (cyl == 6) {
        "background: hsl(230, 70%, 90%); color: hsl(230, 45%, 30%);"
      } else if (cyl == 8) {
        "background: hsl(350, 70%, 90%); color: hsl(350, 45%, 30%);"
      }
      div_out <- htmltools::div(
        style = paste(
          "display: inline-block; padding: 2px 12px; border-radius: 15px; font-weight: 600; font-size: 12px;",
          add_color
          ),
        paste(cyl, "Cylinders")
      )
      
      as.character(div_out) %>% 
        gt::html()
}
```

</details>

<br>

<details>
<summary><strong>Example of a inline bar chart made purely with HTML</summary>  

```{r}
bar_chart <- function(value, color = "red"){
    
    glue::glue("<span style=\"display: inline-block; direction: ltr; border-radius: 4px; padding-right: 2px; background-color: {color}; color: {color}; width: {value}%\"> &nbsp; </span>") %>% 
    as.character() %>% 
    gt::html()
}
```

</details>

<br>

All of these examples can be used in one example table! I've also added a HTML example of a hyperlink for the "data source" which links to the `gt` page for HTML content `r emo::ji("smile")`

```{r}
set.seed(377)
  
mtcars %>% 
  tibble() %>% 
  select(1:4) %>% 
  sample_n(size = 6) %>% 
  mutate(
    rank_change = sample(c("increase", "decrease"), size = 6, replace = TRUE),
    rank_change = map(rank_change, rank_chg)
  ) %>% 
  mutate(
    rating = sample(1:5, size = 6, replace = TRUE),
    rating = map(rating, rating_stars)
    ) %>% 
  mutate(
    cylinder = map(cyl, add_cyl_color)
  ) %>% 
  mutate(
    mpg_plot = mpg/max(mpg) * 100,
    mpg_plot = map(mpg_plot, ~bar_chart(value = .x, color = "lightblue"))
    ) %>% 
  gt() %>% 
  cols_align(
    align = "left",
    columns = vars(mpg_plot)
  ) %>% 
  cols_label(
    mpg = gt::html(as.character(with_tooltip("MPG", "Miles per Gallon")))
  ) %>% 
  tab_source_note(
    source_note = html(
      htmltools::tags$a(
        href = "https://gt.rstudio.com/reference/md.html", 
        target = "_blank", 
        "Data Source"
        ) %>% 
        as.character()
      )
    ) %>% 
  tab_options(
    data_row.padding = px(5)
  )
```

So that's all for now, but hopefully having this "cheatsheet" lets you go even further with all the possible creations you can make with a lot of `gt` and a little bit of HTML!

<br>
<br>