---
title: "Joins vs case whens - speed and memory tradeoffs"
description: |
  Spoiler - joins are fastest, and can be easy to prep!
author:
  - name: Thomas Mock
    url: https://twitter.com/thomas_mock
date: 02-13-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    self_contained: false
    highlight: default
    highlight_downlit: true
    code_folding: false
preview: preview.jpeg
twitter:
  site: "@thomas_mock"
  creator: "@thomas_mock"
categories:
  - tidyverse
  - data cleaning
  - joins
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(datapasta)
library(data.table)
library(bench)
library(espnscrapeR)
library(tictoc)
```

# Matching names

There are many times where you may have a reference dataset that has some type of short-code, abbreviations, nicknames, or even misspellings. In this situation we want to create a new column or even overwrite the existing column with the correct spelling, the full name, etc. 

While there isn't a **wrong** way to do this, I've made mistakes in the past of making things more complicated for myself during data cleaning/prep steps. Most importantly to me in my data cleaning steps are "speed of thought" - aka how long it takes me to type out or reason through, but there are also tradeoffs in memory usage and in execution speed.

I'm going to step through some strategies, honed by years of TidyTuesday data cleaning and prep!

---

In general I want to preface all these examples with the fact that this exercise is easily translatable to various `join` options. Joins are great, but there are situations where the logicals aren't simple matches/index (like `x == "val"`) but rather more complex queries (like `x > 5 & grp == "val"`), but I'm going to live within the constraints of my example use case.

## Example Data

We're going to simulate a dataset with NFL team abbreviations and a "stat" that is relatively in scope with something like EPA, but again this is all simulated data.

```{r}
library(tidyverse)

# relatively nice sized dataset (100,000 rows)
sample_size <- 100000

# Get example team names from espn
all_teams <- espnscrapeR::get_nfl_teams()

# set a reproducible seed (h/t to Anthony's wife's new $4 luggage...)
set.seed(829)

ex_df <- tibble(
  team = sample(all_teams$team_short_name, size = sample_size, replace = TRUE),
  stat = rnorm(sample_size, mean = 0.1, sd = 0.1)
)

ex_df
```

## `dplyr::case_when()` solution

There's a fantastic function in the `dplyr` package called `case_when()`. 

[`case_when`](https://dplyr.tidyverse.org/reference/case_when.html):  

> This function allows you to vectorise multiple if_else()  

In essence, this provides the ability to provide many logical statements to generate a specific outcome. You can read more about the syntax via: `?dplyr::case_when()` in the R console or the [docs](https://dplyr.tidyverse.org/reference/case_when.html). This functions works on vectors as well as in mutations inside data.frames/tibbles/data.tables. Lastly, it can also be translated into SQL automatically via `dbplyr` in situations where you don't want to bring all the data into memory.

`case_when` is very readable to me, and feels better than a whole bundle of nested `ifelse`/`if_else` statements. 

---

A quick `reprex` of `case_when()` usage.

```{r}
ex_car <- mpg %>% 
  select(model, cyl, cty, hwy) %>% 
  slice_sample(n = 100)

ex_car %>% 
  mutate(
    cyl_character = case_when(
      # if logical statement is TRUE then assign new value
      cyl == 4 ~ "Four",
      cyl == 6 ~ "Six",
      cyl == 8 ~ "Eight",
      # This last statement means, if none of the above conditions are met
      # then return a NA value, note that case_when is type-safe
      # so you'll want to use the various NA_??? for the type of output
      TRUE ~ NA_character_
    )
  )
```

---

Now that we understand the syntax a bit better, let's try it on our dataset of interest! Remember, we want to take our team abbreviations and add the team names into the dataset. So we'll need to provide a logical match of each `team` to the correct abbreviation and output the corresponding full name.

This works out very well and we got all the names in the right spot, and with the basic timing via `tictoc` we can see it was still fast. However, there's a LOT of typing, and more importantly a lot of repetitive typing to get this done.

```{r}
library(tictoc)

tic()
ex_df %>% 
    mutate(
      team_name = case_when(
        team == "ARI" ~ "Cardinals",
        team == "ATL" ~ "Falcons",
        team == "BAL" ~ "Ravens",
        team == "BUF" ~ "Bills",
        team == "CAR" ~ "Panthers",
        team == "CHI" ~ "Bears",
        team == "CIN" ~ "Bengals",
        team == "CLE" ~ "Browns",
        team == "DAL" ~ "Cowboys",
        team == "DEN" ~ "Broncos",
        team == "DET" ~ "Lions",
        team == "GB" ~ "Packers",
        team == "HOU" ~ "Texans",
        team == "IND" ~ "Colts",
        team == "JAX" ~ "Jaguars",
        team == "KC" ~ "Chiefs",
        team == "LV" ~ "Raiders",
        team == "LAC" ~ "Chargers",
        team == "LAR" ~ "Rams",
        team == "MIA" ~ "Dolphins",
        team == "MIN" ~ "Vikings",
        team == "NE" ~ "Patriots",
        team == "NO" ~ "Saints",
        team == "NYG" ~ "Giants",
        team == "NYJ" ~ "Jets",
        team == "PHI" ~ "Eagles",
        team == "PIT" ~ "Steelers",
        team == "SF" ~ "49ers",
        team == "SEA" ~ "Seahawks",
        team == "TB" ~ "Buccaneers",
        team == "TEN" ~ "Titans",
        team == "WSH" ~ NA_character_,
        TRUE ~ NA_character_
      )
    )
toc()
```

We can use a few tricks in RStudio to make this easier!

We'll use "multiple cursors" in RStudio and combine that with the amazing [`{datapasta}`](https://github.com/MilesMcBain/datapasta) package from Miles McBain. 

> `datapasta` is about reducing resistance associated with copying and pasting data to and from R. It is a response to the realisation that I often found myself using intermediate programs like Sublime to munge text into suitable formats. Addins and functions in `datapasta` support a wide variety of input and output situations, so it (probably) "just works". Hopefully tools in this package will remove such intermediate steps and associated frustrations from our data slinging workflows.

<aside>

`datapasta` is an absolute life saver for converting objects like vectors/dataframes/clipboard contents and turning them into editable formats in a script.

</aside>

### Step 1: `datapasta`

We'll first create a character vector via `datapasta`, using the below code:

```{r, eval = FALSE}
ex_df %>% 
  distinct(team) %>% 
  arrange(team) %>% 
  pull() %>% 
  datapasta::vector_paste_vertical()
```

I've shown a video of me stepping through this code since there are a lot of different options for `datapasta` and its outputs.

<iframe width="560" height="315" src="https://www.youtube.com/embed/NH3anI8fPjA" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

This gives us a nice starting point of all the vector items and we are ready to work with them in our script.

### Step 2: Multiple line selection in RStudio

I've done this as a video as it's fairly complex to "write about", but very easy to show! The video describes the process of enabling multiple selections in RStudio. 

Note that the enabling multi-line selection is done in RStudio via:  

- Opt key + Left Mouse Button drag on Mac  
- Alt key + Left Mouse Button drag on Windows  

<iframe width="560" height="315" src="https://www.youtube.com/embed/9OgYTCzPqdE" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Once we have this "skeleton" quickly created we can then add the individual "match outputs", like "Cardinals" for "ARI", "Falcons" for "ATL", etc, but we've saved ourself the need to type out a lot of the repeated "skeleton" of the `case_when()`.

## `dplyr` join solution

While the above method is pretty quick to create and understand, we still had to type quite a bit (which is decreased if you use multiple cursors), and in many situations a true join is more efficient in terms of typing, "speed of thought", and execution time.

```{r}
team_join <- all_teams %>% 
  select(team_name, team = team_short_name)

tic()
left_join(ex_df, team_join, by = "team")
toc()
```

Now you may say, well of course I could just use a join if I have a perfect dataset, what do I do with data cleaning?!

We can still "create" a proper joining dataset and use it pretty easily!

### `tribble`

We can use [`tribble`](https://tibble.tidyverse.org/reference/tribble.html) to use almost the _exact_ same setup as the `case_when()`. We can even use the same multiple cursor workflow to take our existing `case_when()` code and turn it into the `tribble` setup.

```{r}
tribble(
  ~team, ~team_name,
  "ARI", "Cardinals",
  "ATL", "Falcons",
  "BAL", "Ravens",
  "BUF", "Bills",
  "CAR", "Panthers",
  "CHI", "Bears",
  "CIN", "Bengals",
  "CLE", "Browns",
  "DAL", "Cowboys",
  "DEN", "Broncos",
  "DET", "Lions",
  "GB", "Packers",
  "HOU", "Texans",
  "IND", "Colts",
  "JAX", "Jaguars",
  "KC", "Chiefs",
  "LV", "Raiders",
  "LAC", "Chargers",
  "LAR", "Rams",
  "MIA", "Dolphins",
  "MIN", "Vikings",
  "NE", "Patriots",
  "NO", "Saints",
  "NYG", "Giants",
  "NYJ", "Jets",
  "PHI", "Eagles",
  "PIT", "Steelers",
  "SF", "49ers",
  "SEA", "Seahawks",
  "TB", "Buccaneers",
  "TEN", "Titans",
  "WSH", NA_character_
  )
```

You could also just create a more traditional `data.frame`/`tibble`, and do the same thing without having to "align" all the rows visually. I _do_ find it a bit easier to look at it row-by-row in the `tribble` so that I can sanity check myself as I go, but up to the reader to decide!

```{r}
ex_df %>% 
  distinct(team) %>% 
  arrange(team) %>% 
  mutate(
    team_name = 
      c(
        "Cardinals", "Falcons", "Ravens", "Bills", "Panthers", "Bears", 
        "Bengals", "Browns", "Cowboys", "Broncos", "Lions", "Packers", "Texans", 
        "Colts", "Jaguars", "Chiefs", "Raiders", "Chargers", "Rams", "Dolphins", 
        "Vikings", "Patriots", "Saints", "Giants", "Jets", "Eagles", "Steelers", 
        "49ers", "Seahawks", "Buccaneers", "Titans", NA_character_
    )
  )
```


## `data.table::fcase`

Now there is also an equivalent function to `case_when()` in `data.table` called [`fcase`](https://github.com/Rdatatable/data.table/issues/3823) (short for fast case when) as of late 2019. Since it operates on any vector, you can use it as a drop in replacement for most in-memory cases of `case_when()` in `dplyr`, or use it in native `data.table` objects. The expectation is that it will operate similarly in output but be more efficient with regards to memory/time.

Note, you do _not_ have the ability to push the translation of `fcase()` into SQL though, as `dbplyr` only has SQL-translation for `case_when()`.

The syntax of `fcase()` and `case_when()` are _essentially_ the same, but rather than a formula-syntax via `~`, it relies on another comma to assign by **position**.

Compare the syntax below:  

```{r}
library(data.table)

mtcars %>% 
  tibble() %>%
  select(mpg, cyl) %>% 
  mutate(
    cyl_case_when = case_when(
      # relies on `~` to indicate assign
      # read as if cyl is equal to 4 assign "Four"
      cyl == 4 ~ "Four",
      cyl == 6 ~ "Six",
      cyl == 8 ~ "Eight",
      TRUE ~ NA_character_
    ),
    cyl_fcase = fcase(
      # relies on another `,` to assign by position
      # reas as if cyl is equal to 4, "Four"
      cyl == 4, "Four",
      cyl == 6, "Six",
      cyl == 8, "Eight",
      TRUE, NA_character_
    ),
    cyl_match = cyl_case_when == cyl_fcase
  )
```

Now, I have converted our example with the `case_when()` statement into `fcase()` for both `dplyr` and `data.table`, but have hidden the code in the detail box below since it's _almost_ identical and would be quite repetitive. I have printed a test to see if the outcomes are equivalent though between native `data.table` and `dplyr`, and they are! `r emo::ji("partying_face")`

Really happy to have the option of using some various methods in-memory via `fcase()` and/or `case_when()` and optionally if working with REALLY large data push some of that compute into the database via `dbplyr` + `case_when()`.

```{r, code_folding = TRUE}
tic()
dplyr_fcase <- ex_df %>% 
    mutate(
      team_name = fcase(
        team == "ARI", "Cardinals",
        team == "ATL", "Falcons",
        team == "BAL", "Ravens",
        team == "BUF", "Bills",
        team == "CAR", "Panthers",
        team == "CHI", "Bears",
        team == "CIN", "Bengals",
        team == "CLE", "Browns",
        team == "DAL", "Cowboys",
        team == "DEN", "Broncos",
        team == "DET", "Lions",
        team == "GB" ,"Packers",
        team == "HOU", "Texans",
        team == "IND", "Colts",
        team == "JAX", "Jaguars",
        team == "KC" ,"Chiefs",
        team == "LV" ,"Raiders",
        team == "LAC", "Chargers",
        team == "LAR", "Rams",
        team == "MIA", "Dolphins",
        team == "MIN", "Vikings",
        team == "NE" ,"Patriots",
        team == "NO" ,"Saints",
        team == "NYG", "Giants",
        team == "NYJ", "Jets",
        team == "PHI", "Eagles",
        team == "PIT", "Steelers",
        team == "SF" ,"49ers",
        team == "SEA", "Seahawks",
        team == "TB" ,"Buccaneers",
        team == "TEN", "Titans",
        team == "WSH", NA_character_,
        TRUE, NA_character_
      )
    )
toc()

# data.table native
dt_fcase <- data.table(ex_df)[, team_name := fcase(
    team == "ARI", "Cardinals",
    team == "ATL", "Falcons",
    team == "BAL", "Ravens",
    team == "BUF", "Bills",
    team == "CAR", "Panthers",
    team == "CHI", "Bears",
    team == "CIN", "Bengals",
    team == "CLE", "Browns",
    team == "DAL", "Cowboys",
    team == "DEN", "Broncos",
    team == "DET", "Lions",
    team == "GB" ,"Packers",
    team == "HOU", "Texans",
    team == "IND", "Colts",
    team == "JAX", "Jaguars",
    team == "KC" ,"Chiefs",
    team == "LV" ,"Raiders",
    team == "LAC", "Chargers",
    team == "LAR", "Rams",
    team == "MIA", "Dolphins",
    team == "MIN", "Vikings",
    team == "NE" ,"Patriots",
    team == "NO" ,"Saints",
    team == "NYG", "Giants",
    team == "NYJ", "Jets",
    team == "PHI", "Eagles",
    team == "PIT", "Steelers",
    team == "SF" ,"49ers",
    team == "SEA", "Seahawks",
    team == "TB" ,"Buccaneers",
    team == "TEN", "Titans",
    team == "WSH", NA_character_,
    TRUE, NA_character_
  )] %>% 
    as_tibble()
```

```{r}
all_equal(dplyr_fcase, dt_fcase)
```

# Speed and Memory

I'm going to preface this by also saying that this is a particular straight forward problem to translate into a `join`, and it has 32 total comparisons. Most uses of `case_when()` or `fcase()` will be more traditional and/or complex logic problems and be much fewer than 30!!!

Now with that being said, you may still say, well I'm not convinced that `join()` method is any easier or faster to code out for humans, and we're all free to make our own decisions! I do want to note that `case_when()` gets memory-inefficient much faster than `???_join()`.

We can really highlight the differences in the execution time and the memory allocated via the `{bench}` package. I'm using `bench` to execute the two different methods 3x time, compare the timing and the memory used, along with some other stats.

In the expandable section below we have a repeat of our above `left_join()` and `case_when()` calls.

```{r, code_folding=TRUE}
join_expr <- function(){
  left_join(ex_df, team_join, by = "team")
}

case_when_expr <- function(){
  ex_df %>% 
    mutate(
      team_name = case_when(
        team == "ARI" ~ "Cardinals",
        team == "ATL" ~ "Falcons",
        team == "BAL" ~ "Ravens",
        team == "BUF" ~ "Bills",
        team == "CAR" ~ "Panthers",
        team == "CHI" ~ "Bears",
        team == "CIN" ~ "Bengals",
        team == "CLE" ~ "Browns",
        team == "DAL" ~ "Cowboys",
        team == "DEN" ~ "Broncos",
        team == "DET" ~ "Lions",
        team == "GB" ~ "Packers",
        team == "HOU" ~ "Texans",
        team == "IND" ~ "Colts",
        team == "JAX" ~ "Jaguars",
        team == "KC" ~ "Chiefs",
        team == "LV" ~ "Raiders",
        team == "LAC" ~ "Chargers",
        team == "LAR" ~ "Rams",
        team == "MIA" ~ "Dolphins",
        team == "MIN" ~ "Vikings",
        team == "NE" ~ "Patriots",
        team == "NO" ~ "Saints",
        team == "NYG" ~ "Giants",
        team == "NYJ" ~ "Jets",
        team == "PHI" ~ "Eagles",
        team == "PIT" ~ "Steelers",
        team == "SF" ~ "49ers",
        team == "SEA" ~ "Seahawks",
        team == "TB" ~ "Buccaneers",
        team == "TEN" ~ "Titans",
        team == "WSH" ~ NA_character_,
        TRUE ~ NA_character_
      )
    )
} 

```

We can then compare their execution multiple times with the `bench` package. This will vary by the execution, but with 3 iterations and 100,000 rows, I have seen about a 10x speed improvement in `left_join` vs `case_when`. Note that in most cases this is still pretty much instantaneous in _"human time"_ as it's all less than 200 ms...

<aside>

Visual reaction time in young adults is in the range of about 250 ms per [Jain et al](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4456887/).

</aside>

However, do note that we use about 28-30 times more memory for the `case_when()` statement.

```{r, warning = FALSE}
bench::mark(
  min_time = 0.1,
  max_iterations = 3,
  min_iterations = 3,
  case_when_expr(),
  join_expr()
) %>% 
  select(expression, min, median, mem_alloc, n_itr)
```

The 150 Mb of memory used, may not seem like a lot with smaller datasets (100,000), but if we were to bump this up to 1,000,000 rows we see a similar 10x increase of memory. Specifically we are now up to ~1.5 Gb of memory used for `case_when()`. So if you're going beyond the millions or 10s of millions of rows AND are trying to do a LOT of comparisons, probably a good idea to start refactoring into a `join` if possible!

```{r, warning=FALSE, code_folding = TRUE}
# 1,000,000 rows as 1e6
sample_size <- 1e6

set.seed(829)

ex_df <- tibble(
  team = sample(all_teams$team_short_name, size = sample_size, replace = TRUE),
  stat = rnorm(sample_size, mean = 0.1, sd = 0.1)
)

bench::mark(
  min_time = 0.1,
  max_iterations = 3,
  min_iterations = 3,
  case_when_expr(),
  join_expr()
) %>% 
  select(expression, min, median, mem_alloc, n_itr) %>% 
  mutate(
    times_faster = 1/(median/max(median)) %>% as.double(),
    mem_used_pct = max(mem_alloc)/mem_alloc
    )
```

There are situations where a `join` doesn't solve the problem, so we can go one step further and add in our `data.table::fcase()` or `dtplyr`-translation of `dplyr` into `data.table`. All the example reprex code is in an expandable chunk below.

```{r, code_folding = TRUE}
library(dtplyr)

case_when_dplyr <- function(){
  ex_df %>% 
    mutate(
      team_name = case_when(
        team == "ARI" ~ "Cardinals",
        team == "ATL" ~ "Falcons",
        team == "BAL" ~ "Ravens",
        team == "BUF" ~ "Bills",
        team == "CAR" ~ "Panthers",
        team == "CHI" ~ "Bears",
        team == "CIN" ~ "Bengals",
        team == "CLE" ~ "Browns",
        team == "DAL" ~ "Cowboys",
        team == "DEN" ~ "Broncos",
        team == "DET" ~ "Lions",
        team == "GB" ~ "Packers",
        team == "HOU" ~ "Texans",
        team == "IND" ~ "Colts",
        team == "JAX" ~ "Jaguars",
        team == "KC" ~ "Chiefs",
        team == "LV" ~ "Raiders",
        team == "LAC" ~ "Chargers",
        team == "LAR" ~ "Rams",
        team == "MIA" ~ "Dolphins",
        team == "MIN" ~ "Vikings",
        team == "NE" ~ "Patriots",
        team == "NO" ~ "Saints",
        team == "NYG" ~ "Giants",
        team == "NYJ" ~ "Jets",
        team == "PHI" ~ "Eagles",
        team == "PIT" ~ "Steelers",
        team == "SF" ~ "49ers",
        team == "SEA" ~ "Seahawks",
        team == "TB" ~ "Buccaneers",
        team == "TEN" ~ "Titans",
        team == "WSH" ~ NA_character_,
        TRUE ~ NA_character_
      )
    )
} 


join_dplyr <- function(){
  left_join(ex_df, team_join, by = "team")
}

join_dtplyr <- function(){
  dt_ex_df <- lazy_dt(ex_df)
  team_join <- lazy_dt(team_join)
    
  dt_ex_df %>% 
    left_join(team_join, by = "team") %>% 
    select(team, stat, team_name) %>% 
    as_tibble() 
}

case_when_dtplyr <- function(){
  lazy_dt(ex_df) %>% 
    mutate(
      team_name = case_when(
        team == "ARI" ~ "Cardinals",
        team == "ATL" ~ "Falcons",
        team == "BAL" ~ "Ravens",
        team == "BUF" ~ "Bills",
        team == "CAR" ~ "Panthers",
        team == "CHI" ~ "Bears",
        team == "CIN" ~ "Bengals",
        team == "CLE" ~ "Browns",
        team == "DAL" ~ "Cowboys",
        team == "DEN" ~ "Broncos",
        team == "DET" ~ "Lions",
        team == "GB" ~ "Packers",
        team == "HOU" ~ "Texans",
        team == "IND" ~ "Colts",
        team == "JAX" ~ "Jaguars",
        team == "KC" ~ "Chiefs",
        team == "LV" ~ "Raiders",
        team == "LAC" ~ "Chargers",
        team == "LAR" ~ "Rams",
        team == "MIA" ~ "Dolphins",
        team == "MIN" ~ "Vikings",
        team == "NE" ~ "Patriots",
        team == "NO" ~ "Saints",
        team == "NYG" ~ "Giants",
        team == "NYJ" ~ "Jets",
        team == "PHI" ~ "Eagles",
        team == "PIT" ~ "Steelers",
        team == "SF" ~ "49ers",
        team == "SEA" ~ "Seahawks",
        team == "TB" ~ "Buccaneers",
        team == "TEN" ~ "Titans",
        team == "WSH" ~ NA_character_,
        TRUE ~ NA_character_
      )
    ) %>% 
    as_tibble()
  
}

fcase_dplyr <- function(){
  ex_df %>% 
    mutate(
      team_name = fcase(
        team == "ARI", "Cardinals",
        team == "ATL", "Falcons",
        team == "BAL", "Ravens",
        team == "BUF", "Bills",
        team == "CAR", "Panthers",
        team == "CHI", "Bears",
        team == "CIN", "Bengals",
        team == "CLE", "Browns",
        team == "DAL", "Cowboys",
        team == "DEN", "Broncos",
        team == "DET", "Lions",
        team == "GB" ,"Packers",
        team == "HOU", "Texans",
        team == "IND", "Colts",
        team == "JAX", "Jaguars",
        team == "KC" ,"Chiefs",
        team == "LV" ,"Raiders",
        team == "LAC", "Chargers",
        team == "LAR", "Rams",
        team == "MIA", "Dolphins",
        team == "MIN", "Vikings",
        team == "NE" ,"Patriots",
        team == "NO" ,"Saints",
        team == "NYG", "Giants",
        team == "NYJ", "Jets",
        team == "PHI", "Eagles",
        team == "PIT", "Steelers",
        team == "SF" ,"49ers",
        team == "SEA", "Seahawks",
        team == "TB" ,"Buccaneers",
        team == "TEN", "Titans",
        team == "WSH", NA_character_
      )
    )
}
  

fcase_dt_native <- function(){
  data.table(ex_df)[, team_name := fcase(
    team == "ARI", "Cardinals",
    team == "ATL", "Falcons",
    team == "BAL", "Ravens",
    team == "BUF", "Bills",
    team == "CAR", "Panthers",
    team == "CHI", "Bears",
    team == "CIN", "Bengals",
    team == "CLE", "Browns",
    team == "DAL", "Cowboys",
    team == "DEN", "Broncos",
    team == "DET", "Lions",
    team == "GB" ,"Packers",
    team == "HOU", "Texans",
    team == "IND", "Colts",
    team == "JAX", "Jaguars",
    team == "KC" ,"Chiefs",
    team == "LV" ,"Raiders",
    team == "LAC", "Chargers",
    team == "LAR", "Rams",
    team == "MIA", "Dolphins",
    team == "MIN", "Vikings",
    team == "NE" ,"Patriots",
    team == "NO" ,"Saints",
    team == "NYG", "Giants",
    team == "NYJ", "Jets",
    team == "PHI", "Eagles",
    team == "PIT", "Steelers",
    team == "SF" ,"49ers",
    team == "SEA", "Seahawks",
    team == "TB" ,"Buccaneers",
    team == "TEN", "Titans",
    team == "WSH", NA_character_
  )] %>% 
    as_tibble()
}

join_dt_native <- function(){
  data.table(ex_df)[data.table(team_join), on = .(team), team_name := team_name] %>% 
    as_tibble()
}
```

Finally we can check the timing/memory usage for all of the combos. In short, `join()`-methods are fastest and use the least memory, `fcase()` whether in native `data.table` or `dplyr` is a bit slower/more memory than `join` but still ~5x faster/more memory efficient than `case_when()`, and `case_when()` is the slowest/most memory hungry (but translates into SQL if needed).

Overall, I'm really happy that we have the ability to mix and match functions from various awesome packages depending on the problems we're trying to solve!

```{r, code_folding = TRUE, warning=FALSE, cache = TRUE}
bench::mark(
  min_time = 0.1,
  max_iterations = 3,
  min_iterations = 3,
  case_when_dplyr(),
  case_when_dtplyr(),
  fcase_dplyr(),
  fcase_dt_native(),
  join_dplyr(),
  join_dt_native(),
  check = FALSE
) %>% 
  select(expression, min, median, mem_alloc, n_itr)
```

